### 2023-06-07 Article.Rmd
literature review - refacut

---
title             : "From Likes to Change: Assessing the Impact of Citizen Engagement on the European Commission's Social Media Platforms"
shorttitle        : "Impact of Citizen Engagement on the European Commission's Social Media Platforms"

author: 
  - name          : "Tanase Tasențe"
    affiliation   : "1"
    corresponding : yes    # Define only one corresponding author
    address       : "Postal address"
    email         : "tanase.tasente@365.univ-ovidius.ro"
    role:         # Contributorship roles (e.g., CRediT, https://casrai.org/credit/)
      - "Conceptualization"
      - "Writing - Original Draft Preparation"
      - "Writing - Review & Editing"
  - name          : "Cristian Opariuc-Dan"
    affiliation   : "1,2"
    role:
      - "Writing - Review & Editing"
      - "Supervision"

affiliation:
  - id            : "1"
    institution   : "Ovidius University, Constanța - Faculty of Law and Administrative Sciences"
  - id            : "2"
    institution   : "Bucharest University"

authornote: |
  Department of Administrative and Social Sciences

abstract: |
  One or two sentences providing a **basic introduction** to the field,  comprehensible to a scientist in any discipline.

  Two to three sentences of **more detailed background**, comprehensible  to scientists in related disciplines.

  One sentence clearly stating the **general problem** being addressed by  this particular study.

  One sentence summarizing the main result (with the words "**here we show**" or their equivalent).

  Two or three sentences explaining what the **main result** reveals in direct comparison to what was thought to be the case previously, or how the  main result adds to previous knowledge.

  One or two sentences to put the results into a more **general context**.

  Two or three sentences to provide a **broader perspective**, readily comprehensible to a scientist in any discipline.

  <!-- https://tinyurl.com/ybremelq -->

keywords          : "keywords"
wordcount         : "X"
bibliography      : ["r-references.bib", "references.bib"]
floatsintext      : no
linenumbers       : yes
draft             : no
mask              : no
figurelist        : no
tablelist         : no
footnotelist      : no
documentclass     : "apa7"
classoption       : "man"
csl               : apa.csl
output            :
  pdf_document    : papaja::apa6_pdf
  word_document   : papaja::apa6_docx
header-includes:
  - \usepackage{ragged2e}
---

```{r setup, include = FALSE}
  library(papaja); library(knitr); library(readxl); library(dplyr); library(psych)
  library(sasLM);library(rstatix); library(naniar); library(lubridate); library(flextable)

  r_refs("r-references.bib")
```

```{r analysis-preferences}
# Seed for random number generation
  set.seed(44542)
  knitr::opts_chunk$set(cache.extra = knitr::rand_seed)
```
# Introduction
\justify
In recent years, the rise of social media platforms has led to significant changes in the way institutions, including public organizations, communicate with their audiences. As a result, the importance of citizen engagement in the process of communication through social media has become increasingly recognized. This engagement involves not only a one-way flow of information from public institutions to citizens, but also an interactive dialogue between the two parties.

The concept of engagement in social media has been studied extensively in recent years, with researchers exploring various aspects of this phenomenon. One such study, conducted by @Dolan2016, approached social media engagement behavior from a uses and gratifications perspective, focusing on the motivations and benefits that users derive from engaging with social media. Meanwhile, the study by @Dragseth2020 explored how social media can be used to build engagement among students in the context of political science education.

Another important aspect of engagement in social media is the role it plays in activation campaigns aimed at consumers. @Mirbagheri2019 conceptualized and developed a scale to measure consumers' engagement with social media activation campaigns. Additionally, @Smith2015 analyzed public engagement with organizations through social media, highlighting the importance of two-way communication between public institutions and citizens.

The differentiating role of platform type in engagement with social media and social media advertising was explored by @Voorveld2018, who found that the level of engagement varies across different social media platforms.

In addition to understanding the various aspects of engagement in social media, it is also important to recognize the significance of citizen engagement in the context of public institutions. Citizen engagement plays a critical role in ensuring transparency and accountability in public decision-making processes. Furthermore, engagement with citizens can lead to the development of more effective policies and programs that better serve the needs of the community.

Moreover, the importance of citizen engagement in the process of communication through social media cannot be overstated. Through social media, public institutions can engage in an interactive dialogue with citizens, build trust, and develop more effective policies and programs. As such, further research, and exploration of the concept of engagement in social media is critical for ensuring that public institutions continue to effectively communicate with and serve the needs of their communities.

Citizen engagement through social media can also contribute to the empowerment of individuals and groups, giving them a voice in public decision-making processes and enabling them to hold public institutions accountable for their actions. This can help to build stronger, more resilient communities that are better equipped to respond to challenges and opportunities.

It is also important to note that while social media has the potential to be a powerful tool for citizen engagement, there are also challenges and risks associated with its use. These include issues related to privacy, security, and the spread of misinformation and disinformation. As such, public institutions must be mindful of these risks and take steps to mitigate them, while also leveraging the power of social media to engage with citizens in a meaningful way.

Overall, the rise of social media has transformed the way public institutions communicate with citizens, placing a greater emphasis on engagement and two-way communication. Understanding the various aspects of engagement in social media is critical for public institutions to effectively communicate with and serve the needs of their communities. By leveraging the power of social media to engage with citizens, public institutions can build trust, empower individuals and groups, and develop more effective policies and programs that better serve the needs of the community.

## Literature review
In studying the impact of citizen engagement on the European Commission's social media platforms, it is crucial to discern the key concepts that underpin this research. These concepts are engagement, social media platforms, and sentiment analysis, which form the backbone of many academic discourses that revolve around these themes.

Engagement, defined as the active interaction of users with digital content, is a fundamental element of any effective social media strategy. It encapsulates various forms of participation, from comments and shares to likes and views [@Mirbagheri2019]. @Smith2015 argue that engagement helps in establishing profound relationships with users, and it can vary according to the type of platform [@Voorveld2018]. The concept has found applicability in numerous contexts, including public health [@heldmanSocialMediaEngagement2013], student learning (Dragseth, 2020), and corporate social responsibility (@doncel-martinCorporateSocialResponsibility2023].

The significance of engagement reaches a higher dimension in the political and public sphere. With social media platforms becoming an integral part of contemporary political communication [@flewPopulismGlobalisationSocial2020; @krzyzanowskiDigitalDiplomacyPolitical2020], citizen engagement has become paramount, as seen in the activities of European Union (EU) agencies (@Muller2022]. There is a myriad of ways in which citizens engage with politics on social media, from commenting on posts to sharing and liking content [@DeWilde2022]. This interaction has been correlated with a spectrum of outcomes, such as influencing voting behavior [@marquartThisTimeNot2020] and attitudes towards vaccination [@mascheriniSocialMediaUse2022].

Simultaneously, social media platforms are becoming recognized as powerful tools for promoting and managing engagement. This dual role is exemplified in the case of the European Commission's activities in Romania [@Rus2021]. Recognizing the potential of these platforms to facilitate citizen engagement, EU institutions have taken steps to optimize their use, implementing various strategies [@Bene2022; @Kanol2021; @ozdemirBirdEyeView2022].

In assessing the impact of citizen engagement on the EU's social media platforms, researchers employ an assortment of methods, one of which is sentiment analysis. This approach, which involves the systematic identification and categorization of the emotional tone behind words, aims to gauge public sentiment, attitudes, and emotions towards specific topics [@Wei2021]. It has been applied in diverse contexts, such as analyzing the emotional distribution in EU smart city communication [@kowalikSocialMediaDistribution2021] and exploring public opinions on climate change policy [@Wei2021].

The European Union’s social media landscape is remarkably complex, as shown by numerous studies. These platforms can simultaneously facilitate positive engagement, such as public service promotion [@Hancu2020], and breed negative phenomena like hate speech [@doncel-martinCorporateSocialResponsibility2023] and digital vigilantism [@allenDigitalVigilantismSocial2020]. Furthermore, the influence of these platforms is shaped by broader societal and political developments, such as migration and smuggling across virtual borders [@Bankston2021].

To sum up, understanding the impact of citizen engagement on the European Commission's social media platforms is a multifaceted issue. These platforms offer opportunities for meaningful citizen engagement and public communication, but their influence is dictated by a complex interplay of individual behavior, institutional strategy, societal trends, and technological developments. This complexity calls for a nuanced understanding of each constituent factor and their collective role in shaping the landscape of citizen engagement within the context of the European Commission's social media platforms. It is a call to researchers, policymakers, and practitioners to continuously explore this evolving realm to maximize the benefits of citizen engagement while mitigating its potential pitfalls.

# Case study
## Research objectives
The primary objective of this research is to conduct a quantitative and qualitative analysis of public engagement across various official European Commission social media platforms (Facebook, Instagram, Twitter, and YouTube). This will involve comparing and contrasting the patterns, trends, and characteristics of online user interactions and responses across these platforms.

## Research method
The research methodology employed in this study involved the extraction and subsequent analysis of data derived from the official Facebook, Twitter, Instagram, and YouTube channels of the European Commission. To amass this data, we utilized the Fanpagekarma platform, a prevalent tool for conducting analytics and monitoring on social media platforms.

The data procured from this platform encompassed a range of elements, including post ID, message content, post type, post date, number of likes, comments, shares, and the rounded figure of followers for each post made by the European Commission in the period from the 3rd of February 2019 to the 23rd of April 2023.

In order to quantify the engagement rate, we calculated the total number of reactions (comprising likes, comments, and shares), and divided it by the total follower count. This engagement rate metric is commonly employed to gauge the extent of audience interaction with a brand or organization on social media platforms. To achieve the research objective, we employed a sentiment analysis method.

## The present study
H1: Communications with emotional resonance are associated with higher public engagement levels.
H2: The effect of emotional resonance on public engagement is moderated by the social media platform utilized.

## Overview of data analysis
The data collected from this research instrument was analyzed using various statistical techniques, including descriptive statistics and ……. Descriptive statistics were used to summarize the data and identify trends and patterns in the engagement rates, likes, comments, and shares for each post type and platform.

In this analysis, we will examine the social media performance of the European Commission's Facebook, Instagram, Twitter, and Youtube accounts during the period from `3 February 2019 to 23 April 2023. We will focus on the engagement rate (ER) of each platform, which measures the percentage of followers who engage with a post by liking, commenting, sharing, or otherwise interacting with it.

We used `r cite_r("r-references.bib")` for all our analyses.

# Results

# Discussion

\newpage

# References

::: {#refs custom-style="Bibliography"}
:::