### 2023-06-13 Article.Rmd
refacut abstract + discussions

---
title             : "From Likes to Change: Assessing the Impact of Citizen Engagement on the European Commission's Social Media Platforms"
shorttitle        : "Impact of Citizen Engagement on the European Commission's Social Media Platforms"
author: 
  - name          : "Tanase Tasențe"
    affiliation   : "1"
    corresponding : yes    # Define only one corresponding author
    address       : "Postal address"
    email         : "tanase.tasente@365.univ-ovidius.ro"
    role:         # Contributorship roles (e.g., CRediT, https://casrai.org/credit/)
      - "Equal contribution"
      - "Conceptualization"
      - "Data curation"
      - "Investigation"
      - "Formal analysis"
      - "Resources"
      - "Software"
      - "Visualization"
      - "Writing - Original Draft Preparation"
      - "Writing - Review & Editing"
  - name          : "Cristian Opariuc-Dan"
    affiliation   : "1,2"
    role:
      - "Equal contribution"
      - "Writing - Original Draft Preparation"
      - "Writing - Review & Editing"
      - "Data curation"
      - "Formal analysis"
      - "Resources"
      - "Software"
      - "Visualization"
  - name          : "Cristina Dana Popescu"
    affiliation   : "1"
    role:
      - "Writing - Original Draft Preparation"
      - "Writing - Review & Editing"
      - "Professional proofreading"
affiliation:
  - id            : "1"
    institution   : "Ovidius University of Constanța, Romania"
  - id            : "2"
    institution   : "University of Bucharest, Romania"
authornote: |
  Department of Administrative and Social Sciences
abstract: |
 This study investigates the role of emotional resonance in social media communications from the European Commission (EC) and its impact on public participation across various platforms including Facebook, Instagram, Twitter, and YouTube. Findings indicate that emotionally resonant communications significantly enhance public engagement, with variations depending on the specific platform. Both positive and negative emotional resonance significantly impact participation, with the latter having a more pronounced effect. Our research highlights the critical importance of platform-specific strategies and the incorporation of emotional resonance in public communications to foster participation in policy dialogues. An analysis of the engagement rates and sentiment across platforms reveals unique tonal variations in the EC's discourse. The effect of emotional resonance on public engagement is also found to be moderated by the choice of social media platform, with certain platforms more significantly moderating the relationship between emotional resonance and engagement than others. This study lays the groundwork for refining communication strategies within a multi-platform digital governance environment.

keywords          : "social media, public institutions, emotional resonance, public engagement, platform-specific strategies, digital governance"
wordcount         : "4067 words in text body 1034 words in reference section"
bibliography      : ["r-references.bib", "references.bib"]
floatsintext      : no
linenumbers       : yes
draft             : no
mask              : no
figurelist        : no
tablelist         : no
footnotelist      : no
documentclass     : "apa7"
classoption       : "man"
csl               : apa.csl
output            :
  pdf_document    : papaja::apa6_pdf
  word_document   : papaja::apa6_docx
header-includes:
  - \usepackage{ragged2e}
---

```{r setup, include = FALSE}
  library(papaja); library(knitr); library(readxl); library(dplyr); library(psych)
  library(sasLM);library(rstatix); library(naniar); library(lubridate); library(flextable)
  library(caret); library(ggpubr); library(lm.beta); library(relaimpo); library(nortest)
  library(interactions); library(misty); library(tm); library(SentimentAnalysis); library(ggplot2)

  r_refs("r-references.bib")
```

```{r analysis-preferences}
# Seed for random number generation
  set.seed(44542)
  knitr::opts_chunk$set(cache.extra = knitr::rand_seed)
```
# Introduction
\justify
In recent years, the rise of social media platforms has led to significant changes in the way institutions, including public organizations, communicate with their audiences. As a result, the importance of citizen engagement in the process of communication through social media has become increasingly recognized. This engagement involves not only a one-way flow of information from public institutions to citizens, but also an interactive dialogue between the two parties.

The concept of engagement in social media has been studied extensively in recent years, with researchers exploring various aspects of this phenomenon. One such study, conducted by @Dolan2016, approached social media engagement behavior from a uses and gratifications perspective, focusing on the motivations and benefits that users derive from engaging with social media. Meanwhile, the study by @Dragseth2020 explored how social media can be used to build engagement among students in the context of political science education.

Another important aspect of engagement in social media is the role it plays in activation campaigns aimed at consumers. @Mirbagheri2019 conceptualized and developed a scale to measure consumers' engagement with social media activation campaigns. Additionally, @Smith2015 analyzed public engagement with organizations through social media, highlighting the importance of two-way communication between public institutions and citizens.

The differentiating role of platform type in engagement with social media and social media advertising was explored by @Voorveld2018, who found that the level of engagement varies across different social media platforms.

In addition to understanding the various aspects of engagement in social media, it is also important to recognize the significance of citizen engagement in the context of public institutions. Citizen engagement plays a critical role in ensuring transparency and accountability in public decision-making processes. Furthermore, engagement with citizens can lead to the development of more effective policies and programs that better serve the needs of the community.

Moreover, the importance of citizen engagement in the process of communication through social media cannot be overstated. Through social media, public institutions can engage in an interactive dialogue with citizens, build trust, and develop more effective policies and programs. As such, further research, and exploration of the concept of engagement in social media is critical for ensuring that public institutions continue to effectively communicate with and serve the needs of their communities.

Citizen engagement through social media can also contribute to the empowerment of individuals and groups, giving them a voice in public decision-making processes and enabling them to hold public institutions accountable for their actions. This can help to build stronger, more resilient communities that are better equipped to respond to challenges and opportunities.

It is also important to note that while social media has the potential to be a powerful tool for citizen engagement, there are also challenges and risks associated with its use. These include issues related to privacy, security, and the spread of misinformation and disinformation. As such, public institutions must be mindful of these risks and take steps to mitigate them, while also leveraging the power of social media to engage with citizens in a meaningful way.

Overall, the rise of social media has transformed the way public institutions communicate with citizens, placing a greater emphasis on engagement and two-way communication. Understanding the various aspects of engagement in social media is critical for public institutions to effectively communicate with and serve the needs of their communities. By leveraging the power of social media to engage with citizens, public institutions can build trust, empower individuals and groups, and develop more effective policies and programs that better serve the needs of the community.

## Literature review
In studying the impact of citizen engagement on the European Commission's social media platforms, it is crucial to discern the key concepts that underpin this research. These concepts are engagement, social media platforms, and sentiment analysis, which form the backbone of many academic discourses that revolve around these themes.

Engagement, defined as the active interaction of users with digital content, is a fundamental element of any effective social media strategy. It encapsulates various forms of participation, from comments and shares to likes and views [@Mirbagheri2019]. @Smith2015 argue that engagement helps in establishing profound relationships with users, and it can vary according to the type of platform [@Voorveld2018]. The concept has found applicability in numerous contexts, including public health [@heldmanSocialMediaEngagement2013], student learning (Dragseth, 2020), and corporate social responsibility (@doncel-martinCorporateSocialResponsibility2023].

The significance of engagement reaches a higher dimension in the political and public sphere. With social media platforms becoming an integral part of contemporary political communication [@flewPopulismGlobalisationSocial2020; @krzyzanowskiDigitalDiplomacyPolitical2020], citizen engagement has become paramount, as seen in the activities of European Union (EU) agencies [@Muller2022]. There is a myriad of ways in which citizens engage with politics on social media, from commenting on posts to sharing and liking content [@DeWilde2022]. This interaction has been correlated with a spectrum of outcomes, such as influencing voting behavior [@marquartThisTimeNot2020] and attitudes towards vaccination [@mascheriniSocialMediaUse2022].

Simultaneously, social media platforms are becoming recognized as powerful tools for promoting and managing engagement. This dual role is exemplified in the case of the European Commission's activities in Romania [@Rus2021]. Recognizing the potential of these platforms to facilitate citizen engagement, EU institutions have taken steps to optimize their use, implementing various strategies [@Bene2022; @Kanol2021; @ozdemirBirdEyeView2022].

In assessing the impact of citizen engagement on the EU's social media platforms, researchers employ an assortment of methods, one of which is sentiment analysis. This approach, which involves the systematic identification and categorization of the emotional tone behind words, aims to gauge public sentiment, attitudes, and emotions towards specific topics [@Wei2021]. It has been applied in diverse contexts, such as analyzing the emotional distribution in EU smart city communication [@kowalikSocialMediaDistribution2021] and exploring public opinions on climate change policy [@Wei2021].

The European Union’s social media landscape is remarkably complex, as shown by numerous studies. These platforms can simultaneously facilitate positive engagement, such as public service promotion [@Hancu2020], and breed negative phenomena like hate speech [@doncel-martinCorporateSocialResponsibility2023] and digital vigilantism [@allenDigitalVigilantismSocial2020]. Furthermore, the influence of these platforms is shaped by broader societal and political developments, such as migration and smuggling across virtual borders [@Bankston2021].

To sum up, understanding the impact of citizen engagement on the European Commission's social media platforms is a multifaceted issue. These platforms offer opportunities for meaningful citizen engagement and public communication, but their influence is dictated by a complex interplay of individual behavior, institutional strategy, societal trends, and technological developments. This complexity calls for a nuanced understanding of each constituent factor and their collective role in shaping the landscape of citizen engagement within the context of the European Commission's social media platforms. It is a call to researchers, policymakers, and practitioners to continuously explore this evolving realm to maximize the benefits of citizen engagement while mitigating its potential pitfalls.

# The present study
Relying on the theoretical framework previously described, the present study aimed to investigate public engagement across various official European Commission social media platforms (Facebook, Instagram, Twitter, and YouTube). We also compared and contrasted the patterns, trends, and characteristics of online user interactions and responses across these platforms. Translating these aims into research questions (RQs), the present study investigated the following:

- *RQ~1~*. How is the communication with emotional resonance associated with higher public engagement levels?
- *RQ~2~*. Does social media platforms influences the emotional resonance on public engagement?
- *RQ~3~*. Does social media platforms moderates the effect of emotional resonance on public engagement?

To answer these questions, we assumed the following:

- *H~1~*: Communications with emotional resonance are associated with higher public engagement levels.
- *H~2~*: Social media platform influences the emotional resonance on public engagement.
- *H~3~*: The effect of emotional resonance on public engagement is moderated by the social media platform utilized.

The novelty of our study lies in the fact that we have systematically compared and contrasted user engagement and emotional resonance across multiple social media platforms utilized by the European Commission. While previous research has generally focused on the impact of citizen engagement within a single social media platform, our study is unique in its comprehensive cross-platform analysis.

In this context, we have pursued an integrated approach to understanding how emotional resonance can influence public engagement levels across these platforms. Furthermore, we have explored the moderating effect of the social media platform itself on the impact of emotional resonance, a novel aspect that has not been fully addressed in previous research.

The innovative approach of this study allowed us to draw robust conclusions about the nuanced relationship between communication strategy, emotional resonance, platform choice, and public engagement. Our research has therefore not only provided key insights into the patterns of citizen engagement on European Commission's social media platforms, but also presented an advanced framework for understanding how such engagement can be effectively harnessed and maximized for public outreach and policymaking.

Ultimately, this study has expanded the current discourse in the field of digital communication and citizen engagement, and opened up new avenues for future research on optimizing communication strategies in a multi-platform social media environment.

# Method
## Procedure used for data gathering
```{r descriptives, echo=FALSE, include=FALSE}
  source("01 - Descriptive analysis.R")
```

We used the Fanpagekarma, a prevalent tool for conducting analytics and monitoring on social media platforms to extract data for the official Facebook, Twitter, Instagram, and YouTube channels of the European Commission. The data included post ID, message content, post type, post date, number of likes, comments, shares, and the rounded figure of followers for each post made by the European Commission, in the period from `r dt.min` to `r dt.max`.

The engagement rate metric is commonly employed to gauge the extent of audience interaction with a brand or organization on social media platforms, and total number of reactions (comprising likes, comments, and shares) were calculated, and divided by the total follower count. A sentiment analysis method was used on engagement rate to reveal the trends and a linear regression analysis was conducted to test the hypothesis.

# Results
## Overview of data analysis
We used `r cite_r("r-references.bib")` for all our analyses.

The initial assumptions assessment was performed by descriptive univariate analysis, data screening for outliers, and missing cases analysis, to verify univariate normality. We further conducted a sentiment analysis, and, finally a linear moderated regression was used for hypothesis testing.

## Preliminary analysis
Some extreme high values were identified on Facebook engagement rate (values over .00484), Instagram engagement rate (values over .0215), Twitter engagement rate (values over .000696) and Youtube engagement rate (values over .000962) and replaced with missing values, however only `r prc.miss`% scores were missing so we decided to remove entire cases.

```{r Descriptives-Table-Ext, echo=FALSE}
   kable(print.univar.row, format = "latex", digits = 8, booktabs = T, caption = "Descriptive analysis. Presence of extreme outliersa")
```

```{r Descriptives-Table-Pur, echo=FALSE}
   kable(print.univar.pur, format = "latex", digits = 8, booktabs = T, caption = "Descriptive analysis. Extreme outliers removed")
```

Results suggested that all engagement rates were highly positively skewed and highly leptokurtic (see Tables \@ref(tab:Descriptives-Table-Ext) and \@ref(tab:Descriptives-Table-Pur)) and the univariate normality assumption of the dependent variable was not met.

\center _Please Insert Tables \@ref(tab:Descriptives-Table-Ext) and \@ref(tab:Descriptives-Table-Pur) around here_ \justify

```{r sentiment, echo=FALSE, include=FALSE}
  source("02 - Sentiment analysis.R")
```
```{r, boxplot-EXT, out.width = '510px', fig.align='center', fig.cap='Boxplot of engagement rates on social media channels. Extremes included', warning=FALSE}
  bp.ext
```
```{r, sent-EXT, out.width = '510px', fig.align='center', fig.cap='Sentiment analysis chart on social media channels. Extremes included', warning=FALSE}
  st.ext
```
```{r, boxplot-NEXT, out.width = '510px', fig.align='center', fig.cap='Boxplot of engagement rates on social media channels. Extremes excluded', warning=FALSE}
  bp.next
```
```{r, sent-NEXT, out.width = '510px', fig.align='center', fig.cap='Sentiment analysis chart on social media channels. Extremes excluded', warning=FALSE}
  st.next
```
## Sentiment analysis
A comprehensive sentiment analysis was conducted on data, both with (see Fig. \@ref(fig:boxplot-EXT) and Fig. \@ref(fig:sent-EXT)) and without extreme outliers (see Fig. \@ref(fig:boxplot-NEXT) and Fig. \@ref(fig:sent-NEXT)) suggesting distinctive tonal variations in the discourse employed by the European Commission across different social media platforms.

\center _Please Insert Figures \@ref(fig:boxplot-EXT) and \@ref(fig:sent-EXT) around here_ \justify

\center _Please Insert Figures \@ref(fig:boxplot-NEXT) and \@ref(fig:sent-NEXT) around here_ \justify

In the case of Facebook, the sentiment's indicator mean value was `r fb.sent.mean.next` (SD=`r fb.sent.sd.next`), showed a tendency to a positive tone (M=`r fb.sent.mean.ext`, SD=`r fb.sent.sd.ext` with extreme outliers). The sentiment ranged between `r fb.sent.min.next` and `r fb.sent.max.next` (`r fb.sent.min.ext` and `r fb.sent.max.ext` with extreme outliers), the discourse with a negative connotation averaged at `r fb.neg.mean.next` (SD=`r fb.neg.sd.next`), while the discourse with a positive undertone demonstrated a mean of `r fb.poz.mean.next` (SD=`r fb.poz.sd.next`), thereby further substantiating a propensity for positivity (M=`r fb.neg.mean.ext`, SD=`r fb.neg.sd.ext`, respectively M=`r fb.poz.mean.ext`, SD=`r fb.poz.sd.ext` with extreme outliers).

In contrast to Facebook, the sentiment indicator on Instagram averaged at `r ig.sent.mean.next`, which points towards a mildly positive sentiment, with the spectrum extending from `r ig.sent.min.next` and `r ig.sent.max.next` (`r ig.sent.min.ext` and `r ig.sent.max.ext` with extreme outliers). The mean sentiment score for negatively perceived discourse was `r ig.neg.mean.next` (SD=`r ig.neg.sd.next`), and the positively perceived discourse showed an average of (M=`r ig.neg.mean.ext`, SD=`r ig.neg.sd.ext`, respectively M=`r ig.poz.mean.ext`, SD=`r fb.poz.sd.ext` with extreme outliers), once again suggesting a bend towards positive discourse on this platform.

Conversely, Twitter displayed a broader sentiment spectrum. Although the sentiment indicator mean was marginally higher than Facebook and Instagram at `r tw.sent.mean.next`, it ranged from `r tw.sent.min.next` and `r tw.sent.max.next` (`r tw.sent.min.ext` and `r tw.sent.max.ext` with extreme outliers), implying a more diverse expression of sentiments. Negative discourse manifested an average of `r tw.neg.mean.next` (SD=`r tw.neg.sd.next`), identical to Instagram but more negative than Facebook. Interestingly, Twitter maintained the highest average score for positive discourse, at `r tw.poz.mean.ext` (SD=`r tw.poz.sd.ext` with extreme outliers), equal to the positivity in Facebook and surpassing that on Instagram.

On YouTube, the sentiment indicator's mean value was observed to be the lowest amongst the evaluated platforms at `r yt.sent.mean.next`, spanning from `r yt.sent.min.next` and `r yt.sent.max.next` (`r yt.sent.min.ext` and `r yt.sent.max.ext` with extreme outliers), thereby indicating a more tempered sentiment. Negative discourse on YouTube scored the lowest mean of `r yt.neg.mean.next` (SD=`r yt.neg.sd.next`), while positive discourse registered the minimum average of `r yt.poz.mean.ext` (SD=`r yt.poz.sd.ext` with extreme outliers) compared to the other platforms.

In conclusion, despite a general inclination towards a slightly positive mean sentiment across all platforms, Twitter demonstrated the most substantial sentiment range, signifying the potential for both intensely negative and positive discourse. Instagram and Facebook depicted a modestly positive sentiment with less variability, while YouTube exhibited the most tempered sentiment range, with the lowest averages for both positive and negative discourse.

```{r prediction, echo=F, include=FALSE}
  source("03 - Regressions.R")
```

A linear regression model was fitted using `r est.n` cases from the purified dataset, values on engagement rate over .000929525 were removed because of extreme outliers and the normality of dependent variable was not met (Anderson-Darling test =`r ad.stat`, p `r ad.p`).

The results suggested that the null hypothesis **H~0~:** *Communications with emotional resonance are not associated with higher public engagement levels* could be rejected (F(`r m1.F.df1`, `r m1.F.df2`)=`r m1.F`, p `r m1.F.p`) and the *H~1~*: *Communications with emotional resonance are associated with higher public engagement levels* was plausible, however the engagement levels was explained only by `r m1.r2.p`% by the positive and negative resonance communications (R^2^=`r m1.r2`, RSR=`r m1.RSE`).

The positive *(B=`r m1.poz.B`, t=`r m1.poz.t`, p  `r m1.poz.p`, \(\beta\)=`r m1.poz.Bt`)* and negative *(B=`r m1.neg.B`, t=`r m1.neg.t`, p  `r m1.neg.p`, \(\beta\)=`r m1.neg.Bt`)* emotional resonance communications were both positively associated statistically significant with engagement rates, and high values on emotional resonance, positive or negative, were associated with high values on engagement rates. However, the relative predictors relevance showed that negative emotional resonance (`r m1.neg.lmg`%) contributed more on engagement rates explanation than positive emotional resonance (`r m1.poz.lmg`%)

Furthermore, we observed that the effect of social media platform was statistically significant on all platforms compared with Twitter and the null hypothesis **H~0~:** *The effect of emotional resonance on public engagement is not moderated by the social media platform* could be rejected. (F(`r m2.F.df1`, `r m2.F.df2`)=`r m2.F`, p `r m2.F.p`). Adding the new as categorical predictor increased the prediction power at `r m2.r2.p`% from `r m1.r2.p`% (R^2^=`r m2.r2`, RSR=`r m2.RSE`), as the most relevant predictor was social network (`r m2.fac.lmg`%), followed by negative emotional resonance (`r m2.neg.lmg`%) and positive emotional resonance (`r m2.poz.lmg`%)

Negative emotional resonance was still statistically significant positively associated by public engagement *(B=`r m2.neg.B`, t=`r m2.neg.t`, p  `r m2.neg.p`, \(\beta\)=`r m2.neg.Bt`)*, but not positive emotional resonance *(B=`r m2.poz.B`, t=`r m2.poz.t`, p  `r m2.poz.p`, \(\beta\)=`r m2.poz.Bt`)*, and compared by Twitter, engagement rates increased statistically significant on Facebook *(B=`r m2.Fcb.B`, t=`r m2.Fcb.t`, p  `r m2.Fcb.p`, \(\beta\)=`r m2.Fcb.Bt`)*, Instagram *(B=`r m2.Isg.B`, t=`r m2.Isg.t`, p  `r m2.Isg.p`, \(\beta\)=`r m2.Isg.Bt`)* and Youtube *(B=`r m2.Ytb.B`, t=`r m2.Ytb.t`, p  `r m2.Ytb.p`, \(\beta\)=`r m2.Ytb.Bt`)*.

Finally, the hypothesis *H~3~*: *The effect of emotional resonance on public engagement is moderated by the social media platform utilized* was also plausible (F(`r m3.F.df1`, `r m3.F.df2`)=`r m3.F`, p `r m3.F.p`), the model with interaction terms explaining `r m3.r2.p`% of engagement rate's variance (R^2^=`r m3.r2`, RSR=`r m3.RSE`) and had a statistically significant better prediction power than the second model (F(`r mod.com.df1`, `r mod.com.df2`)=`r mod.com.F`, p `r mod.com.p`).

The main effect of negative emotional resonance was statistically significant *(B=`r m3.neg.B`, t=`r m3.neg.t`, p  `r m3.neg.p`, \(\beta\)=`r m3.neg.Bt`)*, but not the main effect of positive emotional resonance *(B=`r m3.poz.B`, t=`r m3.poz.t`, p  `r m3.poz.p`, \(\beta\)=`r m3.poz.Bt`)*, and both were moderated by social network. Furthermore, compared to Twitter influence, Facebook *(B=`r m3.Fbk.B`, t=`r m3.Fbk.t`, p  `r m3.Fbk.p`, \(\beta\)=`r m3.Fbk.Bt`)* and Youtube *(B=`r m3.Yut.B`, t=`r m3.Yut.t`, p  `r m3.Yut.p`, \(\beta\)=`r m3.Yut.Bt`)* were statistically significant positively associated with engagement rate, but not Instragram *(B=`r m3.Ins.B`, t=`r m3.Ins.t`, p  `r m3.Ins.p`, \(\beta\)=`r m3.Ins.Bt`)*.

```{r, moderation-1, out.width = '510px', fig.align='center', fig.cap='Interaction effect between social network and negative emotional resonance on engagement rate', warning=FALSE}
  Int.Plot.1
```

The positive association between negative emotional resonance and engagement rate *(B=`r m3.neg.B`, t=`r m3.neg.t`, p  `r m3.neg.p`, \(\beta\)=`r m3.neg.Bt`)* was moderated statistically significant and negatively by Youtube *(B=`r m3.NegYut.B`, t=`r m3.NegYut.t`, p  `r m3.NegYut.p`, \(\beta\)=`r m3.NegYut.Bt`)* and Facebook *(B=`r m3.NegFbk.B`, t=`r m3.NegFbk.t`, p  `r m3.NegFbk.p`, \(\beta\)=`r m3.NegFbk.Bt`)*, messages with negative emotional resonance posted on these social platforms reducing statistically significant the initial positive association. No moderation effect of Instagram was identified on association between negative emotional resonance and engagement rate *(B=`r m3.NegIns.B`, t=`r m3.NegIns.t`, p  `r m3.NegIns.p`, \(\beta\)=`r m3.NegIns.Bt`)* (see Fig. \@ref(fig:moderation-1))

\center _Please Insert Fig. \@ref(fig:moderation-1) around here_ \justify

```{r, moderation-2, out.width = '510px', fig.align='center', fig.cap='Interaction effect between social network and positive emotional resonance on engagement rate', warning=FALSE}
  Int.Plot.2
```

Between positive emotional resonance and engagement rate was no statistically significant association, *(B=`r m3.poz.B`, t=`r m3.poz.t`, p  `r m3.poz.p`, \(\beta\)=`r m3.poz.Bt`)*, but our results showed a positive and statistically significant interaction effect with Instagram *(B=`r m3.PozIns.B`, t=`r m3.PozIns.t`, p  `r m3.PozIns.p`, \(\beta\)=`r m3.PozIns.Bt`)* and Youtube *(B=`r m3.PozYut.B`, t=`r m3.PozYut.t`, p  `r m3.PozYut.p`, \(\beta\)=`r m3.PozYut.Bt`)* (see Fig. \@ref(fig:moderation-2))

\center _Please Insert Fig. \@ref(fig:moderation-2) around here_ \justify

Compared to Twitter posts, if the messages with positive emotional resonance were posted on Instagram or Youtube, the engagement rate increased statistically significant and the effect was most powerfull on Instagram than on Youtube. No interaction effect was observed on Facebook related on positive emotional resonance *(B=`r m3.PozFbk.B`, t=`r m3.PozFbk.t`, p  `r m3.PozFbk.p`, \(\beta\)=`r m3.PozFbk.Bt`)*

# Discussion
The present investigation embarked on a quest to explore public engagement across various social media platforms utilized by the European Commission, with an emphasis on the impact of emotional resonance within content. The role of social media in public institutions' communication strategies is pivotal [@Smith2015], necessitating a deep understanding of engagement dynamics.

The data supports our first hypothesis (H1), suggesting that communications rich in emotional resonance, whether of positive or negative valence, are correlated with enhanced public engagement. However, the ability of emotional resonance to elucidate engagement levels is negligible (0.16%). This corroborates the complex nature of engagement behaviours expounded in scholarly literature [@Dolan2016; @Dragseth2020; @Mirbagheri2019] and points towards the necessity for further examination of supplementary factors influencing public engagement.

Our second hypothesis (H2) posited the influence of the social media platform choice on the effect of emotional resonance on public engagement. The empirical evidence supports this, suggesting that different platforms may cater to varied user expectations and norms. These findings align with prior research that demonstrated platform-specific fluctuations in engagement patterns [@Voorveld2018], underscoring the need for the European Commission to acknowledge heterogeneous engagement patterns and tailor their communications accordingly.

The third hypothesis (H3) asserted that the social media platform would moderate the impact of emotional resonance on public engagement. Our data confirmed this assumption, revealing a statistically significant moderating effect. This suggests that platforms may shape the way the public interacts with emotionally resonant messages, emphasizing the importance of customizing such communications to align with each platform's norms.

Despite these findings, the study is not without limitations. The low R^2 values in our regression models indicate that other unexamined variables might have significant effects on public engagement levels. Future research should persist in exploring these potential factors, acknowledging the complex and multifaceted nature of engagement behaviours.

Additionally, the scope of our research was limited to the European Commission's use of four major social media platforms. Further studies could explore other platforms or public institutions to see if the observed patterns hold. Also, while we focused on emotional resonance, other aspects of communication such as tone, complexity, and framing could be investigated in future research.

In conclusion, this study underscores the significance of emotional resonance and platform-specific norms in driving public engagement. It suggests that forming an emotional connection with the audience through posts could amplify engagement levels. By understanding these dynamics, public institutions like the European Commission can refine their social media strategies, enhance citizen engagement, and bolster the efficacy of their public communication. Despite the limitations, the study provides a valuable starting point for future research into the multifaceted world of social media engagement.

\newpage

# References

::: {#refs custom-style="Bibliography"}
:::